\documentclass[]{article}

%opening
\title{"Double, double toil and trouble; Fire burn and cauldron bubble"\\The Weird Sisters and Global Weirding}
\author{}

\begin{document}

\maketitle


 The word "weird" originally meant \textit{fate}; later it meant \textit{supernatural}, and, more recently, \textit{strange}. In Shakespeare's  \textit{Macbeth}, when the King decides to question the three witches about his fate, he calls them "the weird sisters". So global weirding \textit{might} mean the unfolding of the Earth's fate, or the scientists' efforts to predict the fate; Prof Archer uses it to mean the \textit{strangeness} of what is happening.

 Is it going get hotter? Yes, over most of the planet: that's why they call it Global Warming. Is it going to get colder? Yes, in some places: the North Atlantic between Newfoundland and Ireland ("The Cold Blue Blob") has actually got colder. Last winter was the \textit{coldest on record} there, while the rest of the world had its \textit{hottest year ever}. This was to be expected, as Climate Models predict the ocean currents would weaken, and the Cold Blue Blob is in an area that is warmed by ocean currents. 
 
 Is it going to get dryer? Yes: places that are dry are expected to get dryer. Is it going to get wetter? Yes: places that are wet are predicted to get wetter. Hotter air can hold more water than dry air, so potentially there is more atmospheric moisture to condense as rain. But if the land in one area is too hot, the winds will pass over without discharging their moisture as rain. Instead they will take what little moisture there is, drying the land further.

 Will there be more storms? Some studies suggest that there will be about the same number, but that they will be more intense. This happens because the energy in a storm is locked up as latent heat in water vapour. Higher temperatures mean more moisture, hence more energy. As water vapour rises it cools, releasing energy; higher temperatures mean more energy released. Since storms are chaotic, it is hard to predict what the storm will do.
 
 So global weirding is really about both meanings of the word. Since we've poured so much $CO_{2}$ into the atmosphere, the weird ("fate") of people across the globe will be more weird ("strange") weather. Climate models show that most people live in places with worsening weather, although people in Canada and the North or Russia may benefit, and be able to open ports that work even in tinder. Now that is really weird!
\end{document}
